---
title: "DDSAnalyticsReport"
author: "Quentin,Sita,Olha,Tosin"
date: "4/4/2018"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Abstract

DDS Analytics is a analytics company that specializes in talent management solutions for Fortune 1000 companies. Talent management is defined as the iterative process of developing and retaining employees. It may include workforce planning, employee training programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition). To gain a competitive edge over its competition, DDS Analytics is planning to leverage data science for talent management. The executive leadership has identified predicting employee turnover as its first application of data science for talent management. This repo is the full data analysis of our findings after exploring the many facets of the data.

## I. Introduction





## II. Background



## III.  Methodology

### Data Cleaning & Preparation

```{r}
#a.	Read the csv into R and take a look at the data set.  Output how many rows and columns the data.frame is.
library(readxl)
talentMgmtData <- read_excel("datasets/CaseStudy2-data.xlsx")
dim(talentMgmtData)

#b	The column names are either too much or not enough.  Change the column names so that they do not have spaces, underscores, slashes, and the like. All column names should be under 12 characters. TODO: Make sure you're updating your codebook with information on the tidied data set as well.

names(talentMgmtData) <- c("Age","Attrition","BusinessTrvl","DailyRate","Department","DistFromHome","YrsOfEdu","EduField","EmployeeCnt","EmployeeNum","EnvSatfctn","Gender","HourlyRate","JobInvolmnt","JobLevel","JobRole","JobSatfctn","MaritalStat","MonthlyIncm","MonthlyRate","NumCmpWorked","Over18","OverTime","PrcntSalHike","PerfRating","RlnSatfctn","StandardHrs","StockOptLvl","TtlWrkngYrs","TrngTmsLstYr","WrkLifeBal","YrsAtCompany","YrsInCrntRl","YrsSncLstPrn","YrsWthCurMgr")

names(talentMgmtData)


# c	Some columns are, due to Qualtrics, malfunctioning.

# TODO: Discuss with team, delete Over18 and StandardHrs(80hrs) columns
talentMgmtData$Over18 <- NULL
talentMgmtData <- subset(talentMgmtData, select = -c(StandardHrs))

# d	Make sure your columns are the proper data types (i.e., numeric, character, etc.).  If they are incorrect, convert them. 
str(talentMgmtData)

#TODO: Discuss with team foll. factor levels of dataset
unique(talentMgmtData$BusinessTrvl)

unique(talentMgmtData$JobRole)

unique(talentMgmtData$Attrition)

unique(talentMgmtData$Department)

unique(talentMgmtData$EduField)


library(dplyr)

#Get an Idea of the current format of the colums we are currently using.
glimpse(talentMgmtData)

# Factor appropriate columns 
talentMgmtData$Department <- as.factor(talentMgmtData$Department)
talentMgmtData$BusinessTrvl <- as.factor(talentMgmtData$BusinessTrvl)
talentMgmtData$OverTime <- as.factor(talentMgmtData$OverTime)
talentMgmtData$EduField <- as.factor(talentMgmtData$EduField)
talentMgmtData$Gender <- as.factor(talentMgmtData$Gender)
talentMgmtData$Attrition <- as.factor(talentMgmtData$Attrition)
talentMgmtData$MaritalStat <- as.factor(talentMgmtData$MaritalStat)
talentMgmtData$JobRole <- as.factor(talentMgmtData$JobRole)
talentMgmtData$EnvSatfctn <- as.factor(talentMgmtData$EnvSatfctn)
talentMgmtData$JobLevel <- as.factor(talentMgmtData$JobLevel)
talentMgmtData$StockOptLvl <- as.factor(talentMgmtData$StockOptLvl)
talentMgmtData$PerfRating <- as.factor(talentMgmtData$PerfRating)
talentMgmtData$EmployeeCnt <- as.factor(talentMgmtData$EmployeeCnt)
talentMgmtData$JobInvolmnt <- as.factor(talentMgmtData$JobInvolmnt)
talentMgmtData$RlnSatfctn <- as.factor(talentMgmtData$RlnSatfctn)
talentMgmtData$WrkLifeBal <- as.factor(talentMgmtData$WrkLifeBal)
talentMgmtData$JobSatfctn <- as.factor(talentMgmtData$JobSatfctn)
# handle numeric based columns to show as double or integer if need be.
talentMgmtData$Age <- as.integer(talentMgmtData$Age)
talentMgmtData$EmployeeNum <- as.integer(talentMgmtData$EmployeeNum)
talentMgmtData$DistFromHome <- as.integer(talentMgmtData$DistFromHome)
talentMgmtData$YrsOfEdu <- as.integer(talentMgmtData$YrsOfEdu)
talentMgmtData$DailyRate <- as.numeric(talentMgmtData$DailyRate)
talentMgmtData$TtlWrkngYrs <- as.integer(talentMgmtData$TtlWrkngYrs)
talentMgmtData$YrsAtCompany <- as.integer(talentMgmtData$YrsAtCompany)
talentMgmtData$NumCmpWorked <- as.integer(talentMgmtData$NumCmpWorked)
talentMgmtData$YrsInCrntRl <- as.integer(talentMgmtData$YrsInCrntRl)
talentMgmtData$YrsSncLstPrn <- as.integer(talentMgmtData$YrsSncLstPrn)
talentMgmtData$YrsWthCurMgr <- as.integer(talentMgmtData$YrsWthCurMgr)
talentMgmtData$TrngTmsLstYr <- as.integer(talentMgmtData$TrngTmsLstYr)
#Take a look at data after type conversions are applied to the appropriate columns
glimpse(talentMgmtData)
```


## IV. Results




## IV. Discussion And Conclusions



## References



