---
title: "Understanding Workplace Attrition"
author: "Olha,Sita,Tosin,Quentin"
date: "4/4/2018"
output: 
  pdf_document:
    toc: true
    df_print: kable
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Abstract

DDS Analytics is an analytics company that specializes in talent management solutions for Fortune 1000 companies. Talent management is the iterative process of developing and retaining employees. It may include workforce planning, employee training programs, identifying high-potential employees and reducing/preventing voluntary employee turnover (attrition). To gain a competitive edge over its competition, DDS Analytics is planning to leverage data science for talent management. The executive leadership has identified predicting employee turnover as its first application of data science for talent management. This report is the full data analysis of our findings after exploring the many facets of attrition.

## I. Introduction

Talent is a precious commodity, especially in the corporate sector. Every organization no matter how big or small seeks not only to hire but also to retain the best talent possible. Employee retention, however, is much easier said than done these days, especially when the average amount of years an employee is likely to remain at a specific place of employment is [4.5 years](https://www.forbes.com/sites/davidsturt/2016/01/13/true-or-false-employees-today-only-stay-one-or-two-years/#6736606b6b4c) as of 2014. 

What are the factors that lead to employee attrition? The question of what causes employee attrition is an age-old question that most companies continue to try to answer. Can employee attrition be slowed, or avoided altogether? Is there any data to support that job satisfaction is the primary cause for an employee remaining at a job? All these questions and more will be explored in this study. Our goal is to find sound answers to these questions rooted in a careful statistical analysis of a dataset we have gathered from some of the most influential companies in the world.


## II. Background

As a Talent Management organization, we set out to leverage the strength of Data Science to maintain the best employees at our organization DDS Analytics. For the organization to move forward in this endeavor, it was necessary to study the critical factors in retaining the best employees. As a result, this study was launched by the management team to gain more insight into this arena.


## III.  Methodology

### Data Cleaning & Preparation

To make sure our data was ready for proper analysis we had to go through a series of steps to prepare the data. These steps include the following.

1. Data Import
2. Data Type Conversion for quantitative variables
3. Factorization For Categorical Variables
4. Handling of Missing or Inaccurate Values

We will discuss the process of each step below.

#### Data Import

We have obtained a dataset which includes 35 variables with 1470 observations. We will be referring to the dataset as `talentMgmtData`. To better understand our data we needed to proceed with cleaning it appropriately.

#### Data Type Conversion for quantitative variables

Our analysis is not appropriate if our variables are not showing as the correct data type. Several of the quantitative variables showed up as doubles which were not necessarily fitting for research and analysis. For example, the Number of Training times since last year, and Number Of Company's worked had no reason to show as a double, so it was converted to an integer. We followed this process for each one of the 35 variables to ensure each column assigned data type made sense for the context of the study.

#### Factorization For Catagorical Variables

Within our `talentMgmtData` we noticed several columns that were improperly listed as integers when they are categories. In order for us to be able to do factor analysis we picked out the columns that would be better suited to be categories instead of integers so that we could get a better understanding of how our talent was spread out over different situations. For example, columns like Department, BusinessTravel, Education, & Job satisfaction to name a few, are all better suited to be treated as categorical variables for our research purposes. As a result, we have converted the appropriate columns that explained how our data points are organized by making them factors instead of numerical values.

#### Handling of Missing or Inaccurate Values

For our mathematical calculations to work, we had to convert many of our data points into numerical values so that our regression analysis would be more accurate. For example, DailyRate, Total Working Years, Years with Current Manager, are all instances of numerical based data points that can have mathematical operations performed on them. As a result we located variables like these and made sure the data type was appropriate giving us more flexibility during analysis.

#### DATA CLEANING
```{r echo=FALSE,results='hide',fig.keep='all', warning=FALSE, message=FALSE}
#2a.	Read the csv into R and take a look at the data set.  Output how many rows and columns the data.frame is.
library(readxl)
talentMgmtData <- read_excel("datasets/CaseStudy2-data.xlsx")
dim(talentMgmtData)
names(talentMgmtData)

#2b	The column names are either too much or not enough.  Change the column names so that they do not have spaces, underscores, slashes, and the like. All column names should be under 12 characters. Make sure you're updating your codebook with information on the tidied data set as well.

names(talentMgmtData) <- c("Age","Attrition","BusinessTrvl","DailyRate","Department","DistFromHome","YrsOfEdu","EduField","EmployeeCnt","EmployeeNum","EnvSatfctn","Gender","HourlyRate","JobInvolmnt","JobLevel","JobRole","JobSatfctn","MaritalStat","MonthlyIncm","MonthlyRate","NumCmpWorked","Over18","OverTime","PrcntSalHike","PerfRating","RlnSatfctn","StandardHrs","StockOptLvl","TtlWrkngYrs","TrngTmsLstYr","WrkLifeBal","YrsAtCompany","YrsInCrntRl","YrsSncLstPrn","YrsWthCurMgr")

#changed variable names
names(talentMgmtData)

# 2c	Some columns are, due to Qualtrics, malfunctioning.

# deleting Over18, EmployeeCount, EmployeeNumber and StandardHrs(80hrs) columns that doesn't have much impact on the analysis we are doing

talentMgmtData <- subset(talentMgmtData, select = -c(Over18,EmployeeCnt,EmployeeNum,StandardHrs))

# 2d	Make sure your columns are the proper data types (i.e., numeric, character, etc.).  If they are incorrect, convert them. 

library(dplyr)

#Get an Idea of the current format of the colums we are currently using.
glimpse(talentMgmtData)

#check for NAs
any(is.na(talentMgmtData))

#inspecting the different factor levels of variables of interest
unique(talentMgmtData$BusinessTrvl)
unique(talentMgmtData$JobRole)
unique(talentMgmtData$OverTime)
unique(talentMgmtData$Gender)
unique(talentMgmtData$MaritalStat)
unique(talentMgmtData$Attrition)
unique(talentMgmtData$Department)
unique(talentMgmtData$EduField)

# Factor appropriate columns 
talentMgmtData$Department <- as.factor(talentMgmtData$Department)
talentMgmtData$BusinessTrvl <- as.factor(talentMgmtData$BusinessTrvl)
talentMgmtData$OverTime <- as.factor(talentMgmtData$OverTime)
talentMgmtData$EduField <- as.factor(talentMgmtData$EduField)
talentMgmtData$Gender <- as.factor(talentMgmtData$Gender)
talentMgmtData$Attrition <- as.factor(talentMgmtData$Attrition)
talentMgmtData$MaritalStat <- as.factor(talentMgmtData$MaritalStat)
talentMgmtData$JobRole <- as.factor(talentMgmtData$JobRole)
talentMgmtData$EnvSatfctn <- as.factor(talentMgmtData$EnvSatfctn)
talentMgmtData$JobLevel <- as.factor(talentMgmtData$JobLevel)
talentMgmtData$StockOptLvl <- as.factor(talentMgmtData$StockOptLvl)
talentMgmtData$PerfRating <- as.factor(talentMgmtData$PerfRating)
talentMgmtData$JobInvolmnt <- as.factor(talentMgmtData$JobInvolmnt)
talentMgmtData$RlnSatfctn <- as.factor(talentMgmtData$RlnSatfctn)
talentMgmtData$WrkLifeBal <- as.factor(talentMgmtData$WrkLifeBal)
talentMgmtData$JobSatfctn <- as.factor(talentMgmtData$JobSatfctn)
# handle numeric based columns to show as double or integer if need be.
talentMgmtData$Age <- as.integer(talentMgmtData$Age)
talentMgmtData$DistFromHome <- as.integer(talentMgmtData$DistFromHome)
talentMgmtData$YrsOfEdu <- as.integer(talentMgmtData$YrsOfEdu)
talentMgmtData$DailyRate <- as.numeric(talentMgmtData$DailyRate)
talentMgmtData$TtlWrkngYrs <- as.integer(talentMgmtData$TtlWrkngYrs)
talentMgmtData$YrsAtCompany <- as.integer(talentMgmtData$YrsAtCompany)
talentMgmtData$NumCmpWorked <- as.integer(talentMgmtData$NumCmpWorked)
talentMgmtData$YrsInCrntRl <- as.integer(talentMgmtData$YrsInCrntRl)
talentMgmtData$YrsSncLstPrn <- as.integer(talentMgmtData$YrsSncLstPrn)
talentMgmtData$YrsWthCurMgr <- as.integer(talentMgmtData$YrsWthCurMgr)
talentMgmtData$TrngTmsLstYr <- as.integer(talentMgmtData$TrngTmsLstYr)
#Take a look at data after type conversions are applied to the appropriate columns
glimpse(talentMgmtData)

```

### Preliminary Analysis



---

####3.A####

```{r}
talentData <- talentMgmtData[talentMgmtData$Age > 18, ]
range(talentData$Age)
```

As we can see above, only Ages 19 to 60 exists within our dataset.

Now that the data has been prepared and formatted accordingly it is necessary to explore it to its depths.

####3.B Descriptive statistics####

Now we take a deep dive into our descriptive statistics for some interesting variables in the dataset.
```{r}
library(knitr)
library(kableExtra)
DescriptiveStatistics<-c("min", "max", "mean", "sd")
HourlyRate <- c(min(talentData$HourlyRate), max(talentData$HourlyRate), mean(talentData$HourlyRate), sd(talentData$HourlyRate))
MonthlyIncome <- c(min(talentData$MonthlyIncm), max(talentData$MonthlyIncm), mean(talentData$MonthlyIncm), sd(talentData$MonthlyIncm))
TotalWorkYears <- c(min(talentData$TtlWrkngYrs), max(talentData$TtlWrkngYrs), mean(talentData$TtlWrkngYrs), sd(talentData$TtlWrkngYrs))
Age <- c(min(talentData$Age), max(talentData$Age), mean(talentData$Age), sd(talentData$Age)) 
YearsAtCompany <-c(min(talentData$YrsAtCompany), max(talentData$YrsAtCompany), mean(talentData$YrsAtCompany), sd(talentData$YrsAtCompany))
MonthlyRate <- c(min(talentData$MonthlyRate), max(talentData$MonthlyRate), mean(talentData$MonthlyRate), sd(talentData$MonthlyRate))
YrsOfEdu <- c(min(talentData$YrsOfEdu), max(talentData$YrsOfEdu), mean(talentData$YrsOfEdu), sd(talentData$YrsOfEdu))
table <- data.frame(DescriptiveStatistics, HourlyRate, MonthlyRate, MonthlyIncome, TotalWorkYears, YearsAtCompany, Age,  YrsOfEdu)
table %>%
  kable("latex") %>%
  kable_styling(position= "center", latex_options= "scale_down")
```
The Hourly average rate is $65.88/hour, while the average Monthly Rate is $14,312.21/month. The average Monthly Income is $6,530.21/month, while average Total Worked Years is 11.34 years at a given organization in the dataset. If we examine average Worked years at the company, we see that it is 7.04 years, while the average Age of the employees is 37 years, and average Years of Education of the employees is 2.9 years ('Bachelor' degree).

Below we can see histograms of Hourly Rate and Monthly income.

```{r}
hist(talentData$HourlyRate, col = "darkred", xlab="Hourly Rate", main="Histogram of Hourly Rate")
```

```{r}
hist(talentData$MonthlyIncm, col = "darkgreen", xlab="Monthly Income", main="Histogram of Monthly Income")
```

On the histograms, we can almost see an equal spread of hourly rates within the company, but we can not say the same about monthly income. This visual indicates that employees work a different amount of hours (some of them are part-time, and some of them work with overtime (more than 40hours). We do not have information about whether or not any bonuses are paid in the company, so it does not make sense to continue to analyze working hours). Histogram of income shows a right-skewed distribution of our Monthly Income in our dataset. It is also clear that the majority of the population in this dataset makes between \$1000 and $6000 per month. The higher we go out in income, the more the distribution becomes narrower.

#### 3.C####

##### Understanding Gender, Education, & Occupations

Next, we explore how Gender, Education and Job Role is broken down within our dataset. Below are frequency tables of our findings for these three categories

```{r}
library(ggplot2)

demographics <- talentData[,c("JobRole", "Gender", "EduField")]

build_table <- function(column_name, title) {
  
   t <- as.data.frame(table(column_name))
    names(t) <- c(title, "Frequency")
      t %>% kable("latex") %>% kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive")) 

}

build_table(demographics$Gender, "Gender")

build_table(demographics$EduField, "Education")

build_table(demographics$JobRole, "Job Role")
```


Next, we examine the frequency for gender and how it is distributed across job roles visually.

```{r}

theme_set(theme_light())

ggplot(demographics, aes(demographics$JobRole)) + 
  geom_bar(aes(fill=Gender), width = 0.5) +
  labs(title = "Gender by Job Role",
       subtitle = "Men & Women in Specific JobRole",
       x = "Job Role",
       y = "Frequency") +
  coord_flip()


```

Our chart shows a bimodal distribution of our job roles. What is interesting here is we can see that sales executive is the most commonly occurring job in our dataset with more than 300 people represented in that category. It is also interesting to note that the spread between male and female in that job category looks almost equally represented. Our lowest job role is Human Resources, which is partitioned of mostly men even though there are just under 50 people in this category as a whole.

What about Education? How is education represented across the genders? We take a look at that distribution next.

```{r}

theme_set(theme_dark())

ggplot(demographics, aes(demographics$EduField)) + 
  geom_bar(aes(fill=Gender), width = 0.5) +
  labs(title = "Gender by Field Of Education",
       subtitle = "A breakdown in what fields of education was explored by the genders",
       x = "Field Of Study",
       y = "Frequency") +
  coord_flip()

```

We can see from this chart that Human Resources has the lowest participation regarding education which makes sense given that it is our smallest filled job role. Life Science seems to be the most popular field of study between all the listed education choices with 600 different people in our dataset who studied in this field. This finding does not necessarily match with our discovery regarding our most popular job role. Life Science skills can translate into making great Sales Executives, but this educational field of study does not seem to be directly related to the Sales Executive.

#### 3.D####

Finally, we take a look at a frequency Table for all the different management positions.

```{r}
t <- droplevels(filter(demographics, JobRole == "Manager" | JobRole == "Manufacturing Director" | JobRole == "Research Director"))

build_table(t$JobRole, "ManagementOnly")

```


Now that we have an overall view of our data we can begin to try to answer our inquiries regarding our dataset. Our questions of interest are the following.

1. What are the factors that lead to employee attrition?
2. Can employee attrition be slowed, or avoided altogether?

Our next section will use all of our recent discoveries about the `talentData` to answer these inquiries.

## IV. Deeper Analysis and Visualization


When it comes to jobs the first thing that we want to look at is our age distribution. The understanding of age distribution is a meaningful step in our EDA process as we would like to get an idea of how old or young these individuals are in our entire dataset as it might give us a favorable place to start. We are only interested in exploring individuals in the workforce that are older than 18, so all of the forward analysis will take this constraint into consideration.

```{r}
hist(talentData$Age, xlab= "Ages",
     ylab="Frequency Of Occurrence",
     main= "Age Distribution Frequency",
     col = "blue")
```

Based on the chart above we can see that our age range is between 20 and 60 with a vast portion of ages being between 25 and 45. This is insightful as it might help us with interpretation of our findings moving forward. It is also worth noting that between 30 and 35  is the most occurring age of all the age groups with over 350 recorded ages within this group!

Another thing that we would like to explore is whether or not there is a relationship between our Ages and the number of years they have spent at a specific company. If there is a relationship, we would like to understand whether or not the relationship is positive or negatively correlated.

```{r}
library(ggplot2)
theme_set(theme_light())

ggplot(talentData, aes(x=Age, y=YrsAtCompany)) +
  ggtitle("Years at Company vs Age") +
  theme(plot.title = element_text(hjust = 0.5)) +
  labs(x= "Age", y="Years") +
  geom_point(shape=1, col = "purple") +
  geom_smooth(method = "gam")

```

Our findings here are not surprising. We notice that visually there might be evidence of a positive linear relationship between age and years at a specific company. This finding suggests that the older you are, the more likely you are to have a higher number of years at an individual company. To know for sure, we examine Pearson correlation between `YrsAtCompany` and `Age`.

```{r echo=FALSE, results='hide', warning=FALSE, message=FALSE}
cor.test(x=talentData$Age, y=talentData$YrsAtCompany,
         alternative = "two.sided",
         method="pearson",
         conf.level = 0.95,
         exact = TRUE)
```

Based on the results of our correlation test we have a Pearsons correlation value of `0.302989` (95% CI: 0.25 to 0.34) which is more evidence of a positive linear relationship between Age and Years at a specific company. It is essential for us to keep this relationship in mind moving forward for the rest of the study.

#### Exploring the Relationship between Years At a company and Satisfaction

Now that we know that Age and Income have a positive impact on employee retention we would like to visually confirm our assumption that Job Satisfaction also contributes in a major way to someone remaining at a company. We can do this by examining our different levels of employee satisfaction and their Years they have remained at a particular company.

```{r}
ggplot(talentData, aes(YrsAtCompany)) +
  geom_density(aes(fill=JobSatfctn), alpha=0.8) +
  labs(title= "Years At the Company Density Plot",
       subtitle="Years at the Company grouped by Job Satisfaction",
       x="Years at Company",
       y="Density")


```

The Chart above is very telling. We take a look at our probability distribution across different Job Satisfaction levels. One indicates that there is low employee job satisfaction, and four represents that there is a high employee job satisfaction. If we examine the probability of each within the context of the years an individual stays within a company it becomes clear that lower job satisfaction indicates that this category has the lowest number of years spent at a company. This is no surprise, but it does give us more information regarding negative factors to employee attrition.

##### Correlation Plot of TalentData

In our dataset, there are a large number of numerical based values. We would like to get an idea of the relationship of those variables relates to one another in one glance. Below is a comprised chart of all the correlations regarding our talent data.

```{r}
library(ggcorrplot)

continuous_vars <- talentData[, c("YrsWthCurMgr",
                                  "YrsSncLstPrn",
                                  "YrsInCrntRl",
                                  "YrsAtCompany",
                                  "TrngTmsLstYr",
                                  "TtlWrkngYrs",
                                  "PrcntSalHike",
                                  "MonthlyIncm",
                                  "HourlyRate",
                                  "DailyRate",
                                  "Age",
                                  "DistFromHome",
                                  "YrsOfEdu")]

correlations <- round(cor(continuous_vars), 1)

ggcorrplot(correlations, hc.order = TRUE,
           type = "lower",
           lab = TRUE,
           lab_size = 4,
           method="circle",
           colors= c("tomato2", "white", "springgreen3"),
           title = "Correlation Chart For TalentData",
           ggtheme=theme_bw)
```

The chart above shows us that there are no negative linear relationships in our data between our variables. We can also see that there are some positive relationships between variables that are linearly correlated, and we have now been able to narrow these down, so they are examined in depth.

### 4.C Let's check if there is any relationship between Age and Income. Does Gender make any effect on the Monthly income?

```{r echo=FALSE,results='hide',fig.keep='all'}
MonthlyIncm <- talentData$MonthlyIncm
Age <- talentData$Age
Gender <- talentData$Gender
ggplot(talentData, aes(MonthlyIncm, Age, color = Gender, shape=Gender))+geom_point()+ggtitle("Correlation between Monthly income and Ages")
model_AgeIncome <- lm(MonthlyIncm ~ Age+Gender, data = talentData)
summary(model_AgeIncome)

```
From regression analysis above we can say that Gender does not make significant change in the Monthly employee income. But Age is indeed significant variable (p<0.0001), it can explain 24% of monthly income change.


### 4D. What Factors Cause Employee Turnover?

Our goal is to determine which indicators might lead to employee attrition. The best way for us to find those is to create a regression model for Attrition prediction.

We will now use the stepwise selection method to come up with the model which has only significant variables.

```{r echo=FALSE,results='hide',fig.keep='all'}
library(MASS)
full<- glm(Attrition~Age+BusinessTrvl+DailyRate+Department+DistFromHome+YrsOfEdu+EduField	+EnvSatfctn+Gender+HourlyRate+JobInvolmnt+JobLevel+JobRole+JobSatfctn+MaritalStat+MonthlyIncm+MonthlyRate+ NumCmpWorked+OverTime+PrcntSalHike+PerfRating+RlnSatfctn+StockOptLvl+TtlWrkngYrs+TrngTmsLstYr+WrkLifeBal+	YrsAtCompany+YrsInCrntRl+YrsSncLstPrn+YrsWthCurMgr, family = binomial, data=talentData)
step<- stepAIC(full,trace=FALSE)
step$anova
StepwiseModel_Attrition <- glm(Attrition ~ Age + BusinessTrvl + DailyRate + DistFromHome + EduField + 
    EnvSatfctn + Gender + JobInvolmnt + JobLevel + JobRole + 
    JobSatfctn + NumCmpWorked + OverTime + RlnSatfctn + StockOptLvl + 
    TtlWrkngYrs + TrngTmsLstYr + WrkLifeBal + YrsAtCompany + 
    YrsInCrntRl + YrsSncLstPrn + YrsWthCurMgr, family = binomial, data=talentData)
summary(StepwiseModel_Attrition)
CustomModel_Attrition <- glm(Attrition ~ BusinessTrvl + DistFromHome + 
    EnvSatfctn + JobInvolmnt + 
    JobSatfctn + NumCmpWorked + OverTime + RlnSatfctn + 
    TtlWrkngYrs + WrkLifeBal + YrsAtCompany + 
    YrsInCrntRl + YrsSncLstPrn + YrsWthCurMgr, family = binomial, data=talentData)
summary(CustomModel_Attrition)
```
AIC of the StepwiseModel is 874.95 and R is 1 - (Residual Deviance/Null Deviance)= 1 - 774.95/1285.54 = 0.4. 
AIC of the CustomModel is 996.9 and R is 1 - (Residual Deviance/Null Deviance)= 1 - 944.9/1285.54 = 0.26.

Model assumptions:
Logistic regression does not make many of the assumptions of linear regression and general linear models are based on ordinary least squares algorithms particularly regarding linearity, normality, homoscedasticity, and measurement level.
First, binary logistic regression requires the dependent variable to be binary - Assumption is met.
Second, logistic regression requires the observations to be independent of each other.  In other words, the observations should not come from repeated measurements or matched data - Assumption is met.
Third, logistic regression requires there to be little or no multicollinearity among the independent variables.  This means that the independent variables should not be too highly correlated with each other - Assumption is met.
Fourth, logistic regression assumes linearity of independent variables - Assumption is met.
Finally, logistic regression typically requires a large sample size - Assumption is met (1470 observations).

#####Interpretation of Stepwise Model
We definitely see that Stepwise Model is more predictive. Let's pick the most significant variables that may effect on attrition (positive slope will indicate increasing of attrition chance additively):

- Increasing of Overtime by 1 hour may predict that average attrition possibility will increase by 2.16
- BusinessTrvlTravel_Frequently may predict that average attrition possibility will increase by 2.22
- BusinessTrvlTravel_Rarely may predict that average attrition possibility will increase by 1.19
- Increasing of DistFromHome for 1 mile (assuming that distance were given in miles) may predict that average attrition possibility will increase by 0.06
- If employee has Job role Sales Executive we may predict that average attrition possibility will increase by 1.32
- If employee has Job role Sales Representative we may predict that average attrition possibility will increase by 1.34
- If number of companies where an employee worked increases by 1, we may predict that average attrition possibility will increase by 0.2
- Increasing of Years after last Promotion by 1 year may predict that average attrition possibility will increase by 0.16

```{r}
variables<-c("Overtime","BusinessTrvlTravel_Frequently", "BusinessTrvlTravel_Rarely", "DistFromHome", "NumCmpWorked", "YrsSncLstPrn")
Effect<-c(2.16, 2.22, 1.19, 0.06, 0.2,0.16)
df <- data.frame(variables,Effect)
mytheme <- theme(plot.title = element_text(face = "bold.italic",size= 14, color = "black"), axis.text.x = element_text(face ="bold.italic"), axis.text.y = element_text(face = "bold.italic", size = 7))
df$variables <- factor(variables, levels = df$variables[order(df$Effect)])
ggplot(data = df, aes(x = df$variables, y = df$Effect), horis=TRUE)+geom_bar(stat = "identity", fill="#FF9999")+labs(title = "Attrition influential factors", x="Variables",y="Influential Percent")+mytheme+coord_flip()
```

As we can see from the histogram above, the most influential factors for Attrition are Business Travels and Overtime. Let's see Overtime and Business Travels by Attrition factors on histograms and confirm this conclusion with actual percentage.

```{r}
ggplot(talentData,aes(x=OverTime,fill = Attrition))+geom_bar()+ggtitle("Attrition based on Overtime factors")  +labs(x = "OverTime", y = "Count")
library(kableExtra)
mytable1 <- xtabs(~OverTime+Attrition, data = talentData)
tabl1<- data.frame(ftable(mytable1))
percent11<- tabl1[3,3]/(tabl1[3,3]+tabl1[1,3])*100
percent12<- tabl1[4,3]/(tabl1[4,3]+tabl1[2,3])*100
Attrition_Percent <-round(c(0,0,percent11,percent12),2)
df1 <- data.frame(tabl1,Attrition_Percent)
df1 %>%
  kable("latex") %>%
  kable_styling() %>%
  row_spec(4, bold = T, color = "white", background = "#D7261E")
```
Attrition rate within employees with OverTime is 30.19%. It is 193% higher then attrition rate within employees without Overtime working hours.

```{r}
ggplot(talentData,aes(x=BusinessTrvl,fill = Attrition))+geom_bar()+ggtitle("Attrition based on Business Travel factors")  +labs(x = "Business Travel", y = "Count")
mytable2 <- xtabs(~BusinessTrvl+Attrition, data = talentData)
tabl2<- data.frame(ftable(mytable2))
percent21<- tabl2[4,3]/(tabl2[4,3]+tabl2[1,3])*100
percent22<- tabl2[5,3]/(tabl2[5,3]+tabl2[2,3])*100
percent23<- tabl2[6,3]/(tabl2[6,3]+tabl2[3,3])*100
Attrition_Percent <-round(c(0,0,0,percent21,percent22,percent23),2)
df2 <- data.frame(tabl2,Attrition_Percent)
df2 %>%
  kable("latex") %>%
  kable_styling() %>%
  row_spec(5:6, bold = T, color = "white", background = "#D7261E")
```
Attrition rate within employees who has Frequent Business Travels is 24.36%, and 14.89% for those who has Rarely Business Travels. It is 223% and 97.7% higher then the attrition rate within NON travel employees, in respect to Frequent and Rarely Business Travels.

Please see below a histogram with variables that may have a good affect for "Long stay" employees, those who are satisfied with their working position and do not want to quit.

```{r}
variables<-c("EnvSatfctn2","EnvSatfctn3", "EnvSatfctn4", "JobInvolmnt2", "JobInvolmnt3", "JobInvolmnt4",
             "JobSatfctn3","JobSatfctn4", "RlnSatfctn2", "RlnSatfctn3", "RlnSatfctn4", "StockOptLvl1",
             "StockOptLvl2","WrkLifeBal3", "YrsWthCurMgr", "TrngTmsLstYr")
Effect<-c(-1.10, -1.24, -1.37,  -1.19, -1.54,-2.11, -0.67, -1.29,  -0.98,  -1.03,  -1.03,  -1.46,  -1.45, -1.5, -0.14, -0.17)
df <- data.frame(variables,Effect)
mytheme <- theme(plot.title = element_text(face = "bold.italic",size= 14, color = "black"), axis.text.x = element_text(face ="bold.italic"), axis.text.y = element_text(face = "bold.italic", size = 7))
df$variables <- factor(variables, levels = df$variables[order(df$Effect)])
ggplot(data = df, aes(x = df$variables, y = df$Effect), horis=TRUE)+geom_bar(stat = "identity", fill="#66CC99")+labs(title = "'Long stay' influential factors", x="Variables",y="Influential Percent")+mytheme+coord_flip()
```

Lets confirm our findings with actual numbers.

```{r}
ggplot(talentData,aes(x=JobInvolmnt,fill = Attrition))+geom_bar()+ggtitle("Attrition based on Job Involvement factors") +labs(x = "Job Involvement level", y = "Count")
mytable3 <- xtabs(~JobInvolmnt+Attrition, data = talentData)
tabl3<- data.frame(ftable(mytable3))
percent31<- tabl3[5,3]/(tabl3[5,3]+tabl3[1,3])*100
percent32<- tabl3[6,3]/(tabl3[6,3]+tabl3[2,3])*100
percent33<- tabl3[7,3]/(tabl3[7,3]+tabl3[3,3])*100
percent34<- tabl3[8,3]/(tabl3[8,3]+tabl3[4,3])*100
Attrition_Percent <-round(c(0,0,0,0,percent31,percent32,percent33,percent34),2)
df3 <- data.frame(tabl3,Attrition_Percent)
df3 %>%
  kable("latex") %>%
  kable_styling() %>%
  row_spec(7:8, bold = T, color = "white", background = "#66CC99")
```
Average Attrition rate for those who has Job Involvement level "high" and "very high" is smaller by 41.43% comparing with those who has Job Involvement level "low" and "medium". 

```{r}
ggplot(talentData,aes(x=StockOptLvl,fill = Attrition))+geom_bar()+ggtitle("Attrition based on Stock Option factors") +labs(x = "Stock Option Level", y = "Count")
mytable4 <- xtabs(~StockOptLvl+Attrition, data = talentData)
tabl4<- data.frame(ftable(mytable4))
tabl4
percent41<- tabl4[5,3]/(tabl4[5,3]+tabl4[1,3])*100
percent42<- tabl4[6,3]/(tabl4[6,3]+tabl4[2,3])*100
percent43<- tabl4[7,3]/(tabl4[3,3]+tabl4[7,3])*100
percent44<- tabl4[8,3]/(tabl4[8,3]+tabl4[4,3])*100
Attrition_Percent <-round(c(0,0,0,0,percent41,percent42,percent43,percent44),2)
df4 <- data.frame(tabl4,Attrition_Percent)
df4 %>%
  kable("latex") %>%
  kable_styling() %>%
  row_spec(6:7, bold = T, color = "white", background = "#66CC99")
```
It is very interesting that employees prefer Stock Opt Level 1 and 2 verses Stock Opt Level 0 and 3. Average attrition rate is smaller by 61.4% for those who has Stock Opt Level 1 and 2.

## V. Discussion And Conclusions

After extensive statistical analysis, we have been able to conclude that there are indeed several factors that lead to the attrition of employees. Based on what these factors are there are things the business can do to make it much harder for someone to decide if they want to leave a company. We have compiled a list of recommendations that DDSAnalytics can follow to help reduce Attrition. It should be stated that the following factors below will not guarantee employee Attrition will stop, but it will increase the chances of slowing Attrition at an organization.

#### Limit Business Travel

We found that Frequent business travel was a considerable contributor to churn in the dataset. If a company wants to reduce the chances of someone wanting to leave a specific job they should make sure that the worker has limited responsibility to travel as it could make all the difference in deciding whether to change companies.

#### Understand How workers feel about Overtime

We found that not having overtime was very significant in determining employee attrition. Too much overtime might indicate that management should limit the number of times they ask employees to pick up extra hours, or only consider asking those that need the extra hours for personal reasons whether or not they want to work more. Poorly managing overtime can go a long way into contributing to employee attrition.

#### Consider Remote Work

Based on studying the significance of the distance traveled from home to get to work, i.e., commuting we found that this is an immensely significant factor. Based on a 95% confidence Interval the optimal distance is somewhere between 2 miles to 6 miles. If an employee has to travel more than that one should consider offering a remote option in order to limit attrition due to the Distance from Home.

#### Consider Hiring people that worked for less number of Companies

After studying the data, we found that the ideal number of companies worked for those that did experience attrition was 1 to 3 companies on a 95% confidence interval.  High unsettled employee numbers indicate that people that work for multiple companies are more likely to contribute to the Attrition rate. This proposal is an extreme solution, but one should consider the number of companies a potential employee worked before hiring them, as it might be a key predictor as to whether or not that individual will stay or leave.

#### Satisfaction Indicators

We found that Job Involvement Level 4 ("very high") and 3 ("high"), Environment Satisfaction, Job Satisfaction Level 4 ("very high"), Stock Option Levels 1 and 2 are all contributed to employee churn in a major way. This indicates that management should consider doing things to make the work environment more comfortable so that the employees feel great about going to work every day. Employees also need to feel heavily involved in their job in as well as be satisfied doing it. It is important to conduct reviews and surveys to find out how employees feel in these areas so that an overall general pulse can be gathered on how the company is performing here.

#### Work relationships must be maintained

Employees consider work relationships to be a key factor and it also contributes to employee "long stay" position. It makes sense to have relationship building events that improve ties within employee to employee relationships. People have to like who they work with, and management can improve this by tracking how their employees are relating to each other on the job. Management should consider bettering situations that cause bad relationships between workers.

## References

[Forbes "True Or False Employees Today only stay one or Two Years : David Sturt and Todd Nordstrom "](https://www.forbes.com/sites/davidsturt/2016/01/13/true-or-false-employees-today-only-stay-one-or-two-years/#6736606b6b4c)

